---
title: "Covid91 vaccine study Final2024"
author: "Roselyn Villator"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Fall 2024" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

##Methods

Question: Is the covid91 vaccine effective in males?

$H_0$: The vaccine is not effective in preventing covid91 in males.

$H_A$: The vaccine if effective in preventing covid91 in males.

infected~sex(M)
categorical~categorical

We'll be using Cat~Cat

## Graphical Results

```{r}
Male <- FinalDat[FinalDat$sex=='M',]
```

```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected + treatment,data=Male, type="percent")

```

From this bar chart it shows that you are less likely to contract covid91 if you did receive the vaccine.

## Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows the results in percentages. Within the placebo group 73.80% of individuals got infected with covid91. While in the vaccine group 26.20% contracted covid91. This indicates that receiving the vaccine as a male significantly reduces the risk of contracted covid91.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

This data 
## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions