---
title: "Covid91 vaccine study Final2024"
author: "Roselyn Villator"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Fall 2024" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

##Methods

Question: Is the covid91 vaccine effective in males?

$H_0$: The vaccine is not effective in preventing covid91 in males.

$H_A$: The vaccine is effective in preventing covid91 in males.

infected~sex(M)
categorical~categorical

We'll be using Cat~Cat

## Graphical Results

```{r}
Male <- FinalDat[FinalDat$sex=='M',]
```

```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected + treatment,data=Male, type="percent")

```

From this bar chart it shows that you are less likely to contract covid91 if you did receive the vaccine.

## Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows the results in percentages. Within the placebo group 73.80% of individuals got infected with covid91. While in the vaccine group 26.20% contracted covid91. This indicates that receiving the vaccine as a male significantly reduces the risk of contracted covid91.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

This data shows that the Chi-Square P-Value is 2.2e-16 and the Fisher exact test shows that the P-value is 2.2e-16 which is significantly lower than 0.05 which means we reject the null hypothesis.

## Conclusion

The bar chart shows that you are less likely to contract covid91 if you did receive the vaccine. The numerical data shows the results in percentages. Within the placebo group 73.80% of individuals got infected with covid91. While in the vaccine group 26.20% contracted covid91. This indicates that reveiving the vaccine as a ale significantly reduces the risk of contracted covid91. The inferential data shows a Chi-Square P-Value of 2.2e-16, this also includes the Fisher exact test shows that the P-value is 2.2e-16 which is significantly lower than 0.05 which means we reject the null hypothesis. All in all these examples and simulations show that as a male you are far less likely to contract covid91 upon receiving the vaccine.

## Females

## Methods 

Question: Is the vaccine effective in preventing covid91 in females?

$H_0$: The vaccine is not effective in preventing covid91 in females.

$H_A$: The vaccine is effective in preventing covid91 in females.

infected~sex(F)
categorical~categorical

We'll be using Cat~Cat

## Graphical Results

```{r}
Female <- FinalDat[FinalDat$sex=='F',]
```

```{r}
barchartGC(~infected + treatment,data=Female)
barchartGC(~infected + treatment,data=Female, type="percent")

```

This bar chart shows that females are less likely to contract covid91 after reveiving the vaccine.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Female)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows the results in percentages. Within the placebo group 64.93% of individuals got infected with covid91. While in the vaccine group 35.07% contracted covid91. This indicates that receiving the vaccine as a female significantly reduces the risk of contracted covid91.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi-Square test shows a P-value of 1.914e-08 and the fisher exact test shiws that the P-value is 1.246e-08 which is significantly lower than 0.05 which means we reject the null hypothesis.

## Conclusion

The bar chart shows that uou are less likely to contract covid91 if you did reveive the vaccine. The numerical data shows the results in percentages. Within the placebo group 64.93% of indivuals got infected with covid91. While in the vaccine group 35.07% contracted covid91. This indicates that receiving the vaccine as a female reduces the risk of contracted covid91. The inferential data shows a Chi-square p-value of 1.914e-08, and this data also includes the Fisher exact test shows that the p-value is 1.246e-08 which is significantly lower than 0.05 which means we reject the null hypothesis. In summation these examples and simulations shows that as a female you are far less likely to contract covid91 if you receive the vaccine.

## LGBTQ

## Methods

Question: Is the vaccine effective in preventing covid91 in members of the LGBTQ community?

$H_0$: The vaccine is not effective in preventing covid91 in members of the LGBTQ community.

$H_A$: The vaccine is effective in preventing covid91 in members of the LGBTQ community.

infected~LGBTQ(gay)
categorical~categorical

We'll be using Cat~Cat

## Graphical Results

```{r}
LGBTQ <- FinalDat[FinalDat$LGBTQ=='gay',]
```

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

This bar chart shows that members of the LGBTQ community are less likely to contract covid91 after reviewing the vaccine.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows the results in percentages. Within the placebo group 80.00% of individuals got infected with covid91. While in the vaccine group 20.00% contracted covid91. This indicates that reveiving the vaccine as a LGBTQ community member significantly reduces the risk of contracted covid91.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi-square test shows that the p-value is 2.266e-06 and data the fisher exact test shows that the p-value is 7.269e-07 which is significantly lower than 0.05 which means we reject the null hypothesis.

## Conclusion
The bar chart shows that you are less likely to contract covid91 if you did receive the vaccine. The numerical data shows the results in percentages. Within the placebo group 80.00% of indivuals got infected with covid91. While in the vaccine group 20.00% contracted covid91. This indicates that recieving the vaccine as a LGBTQ member significantly reduces the risk of contracted covid91. The inferential data shows chi-squared p-value of 2.266e-06, this data also includes the fisher exact test that shows that the p-value is 7.269e-07 which is significantly lower than 0.05 which means we reject the null hypothesis. In summation these examples and simulations show that as a member of the LGBTQ community you are far less likely to contact covid91 if you receive the vaccine.

## Druggies

## Methods

Question: Is the vaccine effective in preventing covid91 in a drug user?

$H_0$: The vaccine is not effective in preventing covid91 in  drug users.

$H_A$: The vaccine is effective in preventing covid91 in drug users.

infected~DrugUser(yes)
categorical~categorical

We'll be using Cat~Cat

## Graphical Results

```{r}
DrugUser <- FinalDat[FinalDat$DrugUser=='yes',]
```

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```

According to the bar chart you are more likely to get covid91 if recieve the vaccine as a drug user.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows the results in percentages. Within the placebo group 8.33% of individuals got infected with covid91. While in the vaccine group 91.67% contracted covid91. This indicates that receiving the vaccine as a drug user significantly increases the risk of contracted covid91.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-square test shows a p-value of 1.414e-10 data and the fisher exact test shows that the p-value is 3.261e-12 which is significantly lower than 0.05 which means we reject the null hypothesis.

## Conclusion

The bar chart shows that you are far more likely to contract Covid91 if you did receive the vaccine as a drug user. The numerical data shows the results in percentages. Within the placebo group 8.33% of individuals got infected with Covid91. While in the vaccine group 91.67% contracted covid-91.This indicates that receiving the vaccine as a drug user significantly increases the risk of contracting Covid-91. The inferential data shows a Chi-Squared P-Value of 1.414e-10, and the data includes a Fisher exact test showing that the P-Value is 3.261e-12 which is significantly lower than 0.05 which means we reject the null hypothesis.In summation these examples and simulations show that as a drug user you are far more likely to contract covid-91 if you receive the vaccine.

# Overall Results and Conclusions